---
title: "STAA 566 HW3"
author: "Ryan Miller"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r, echo = FALSE, message = FALSE}
## Load required libraries
library( gridExtra  )
library(  mapproj   )   # Req'd for mapping                 TODO: Remember to uninstall
library(   maps     )   # Req'd for mapping                 TODO: Remember to uninstall
library(    sf      )   # Req'd for simple features mapping TODO: Remember to uninstall
library( tidycensus )   # Req'd for data retrieval          TODO: Remember to uninstall
library( tidyverse  )
library(   units    )   # Req'd for mapping                 TODO: Remember to uninstall
library(  viridis   )
# TODO: Also remember to uninstall libudunits2-dev from computer
# TODO: Also remember to uninstall libgdal-dev     from computer

## Clear function for ease (by default keeps all functions)
# Optional input E - character array of exceptions
# Optional input P - character array of inclusions
clear <- function( all = FALSE, E = "", P = "" )
{
    # By default don't clear the loaded data unless intentionally cleared
    if( ! all ){ E <- c( E, 'api.key' ) }

    # Get list of functions and exclude anything in E
    l <- setdiff( ls(envir=parent.frame()), c( lsf.str(envir=parent.frame()), E ) )

    # Add P to the list.
    if( all( P != "" ) ){ l <- c( l, P ) }

    # Remove the stuff
    rm( list = l, envir = parent.frame() )
}

## Clear things
clear( )

##########################################################################################
## End Setup
##########################################################################################

```
\
\
The data for this project comes from the US Census Boundary and Attribute Data via
[tidycensus](https://cran.r-project.org/web/packages/tidycensus/index.html).

The maps display the proportion of one-person, two-person, three-person, and
four-plus-person households by state in the United States of America.

I think that this data gives insight to the life-styles in different parts of the country,
and therefore I find it interesting.

I chose to display the maps using `coord_map( 'conic', lat0 = 30 )` in order for the map
to not be skewed; i.e., the sizes of the states to not be distorted. I used the
`viridis` color package to color the maps with the corrected color map `cividis`.

```{r, echo = FALSE, message = FALSE}
## Get CENSUS_API_KEY for querying data
if( !exists( 'api.key' ) || api.key == '' )
{
    readRenviron( '~/.Renviron' )
    api.key <- Sys.getenv( 'CENSUS_API_KEY' )
}
if( api.key == '' ){ stop('API KEY ERROR'); }

## Get the household size
hs <- get_decennial( geography = 'state', variables = c( 'H013001', 'H013002', 'H013003',
                                                        'H013004', 'H013005', 'H013006',
                                                        'H013007', 'H013008' ) )

## Get U.S.A state data
states <- map_data( 'state' ) %>% mutate( region = str_to_title( region ) )
```

```{r, echo = FALSE, results = 'hide'}
head( hs )
head( states )

## There's almost certainly an easier way to do this in tidyverse
tot   <- which( hs$variable == 'H013001' )
op    <- which( hs$variable == 'H013002' )
twp   <- which( hs$variable == 'H013003' )
thp   <- which( hs$variable == 'H013004' )
frp   <- which( hs$variable == 'H013005' )
fvp   <- which( hs$variable == 'H013006' )
sxp   <- which( hs$variable == 'H013007' )
spp   <- which( hs$variable == 'H013008' )
p.op  <- hs$value[ op ] / hs$value[ tot ]
p.twp <- hs$value[ twp ] / hs$value[ tot ]
p.thp <- hs$value[ thp ] / hs$value[ tot ]
p.fpp <- sum( c(hs$value[ frp ], hs$value[ fvp ], hs$value[ sxp ], hs$value[ spp ] ) ) / hs$value[ tot ]

## Update the housing data
h.data <- hs[ tot, ]
h.data <- h.data %>% select( -c(variable, value ) ) %>%
    mutate( Total = hs$value[ tot ],
    One.Person = hs$value[ op ] / Total,
    Two.Person    = hs$value[ twp ] / Total,
    Three.Person  = hs$value[ thp ] / Total,
    Four.Person   = hs$value[ frp ] / Total,
    Five.Person   = hs$value[ fvp ] / Total,
    Six.Person    = hs$value[ sxp ] / Total,
    SevenP.Person = hs$value[ spp ] / Total )

# Ensure percentages sum to 1
rowSums( data.matrix( h.data[ , 4:10 ]) )

## Plot data for the four plots
pd <- vector( 'list', 8 )
pd[[1]] <- h.data %>% mutate( p = One.Person    ) %>% select( NAME, p )
pd[[2]] <- h.data %>% mutate( p = Two.Person    ) %>% select( NAME, p )
pd[[3]] <- h.data %>% mutate( p = Three.Person  ) %>% select( NAME, p )
pd[[4]] <- h.data %>% mutate( p = Four.Person   ) %>% select( NAME, p )
pd[[5]] <- h.data %>% mutate( p = Five.Person   ) %>% select( NAME, p )
pd[[6]] <- h.data %>% mutate( p = Six.Person    ) %>% select( NAME, p )
pd[[7]] <- h.data %>% mutate( p = SevenP.Person ) %>% select( NAME, p )
pd[[8]] <- h.data %>% mutate( p = Six.Person + SevenP.Person ) %>% select( NAME, p )

pd[[1]]$title <- 'One Person Households'
pd[[2]]$title <- 'Two Person Households'
pd[[3]]$title <- 'Three Person Households'
pd[[4]]$title <- 'Four Person Households'
pd[[5]]$title <- 'Five Person Households'
pd[[6]]$title <- 'Six Person Households'
pd[[7]]$title <- 'Seven Person Households'
pd[[8]]$title <- 'Six+ Person Households'

## Get a map of the U.S.A
h <- lapply( pd, function( p )
            {
                ## TODO: Figure out how to do consistent color bar across all
                pd_ <- states %>% left_join( p, c( 'region' = 'NAME' ) )
                ttl <- p$title[1]

                h_ <- ggplot( data = pd_, aes( x = long, y = lat, group = group,
                                              fill = p ) ) +
                    geom_polygon( color = 'white' ) +
                    scale_fill_viridis( option = 'cividis', direction = -1 ) +
                    theme_minimal() + theme( axis.title = element_blank() ) +
                    ggtitle( ttl ) +
                    guides( fill = guide_legend( title = 'Percent' ) ) +
                    coord_map( 'conic', lat0 = 30 )

                return( h_ )
            } )
```

```{r, echo = FALSE, fig.align = 'center', results = 'hide'}
grid.arrange( h[[1]], h[[2]], h[[3]], h[[4]], h[[5]], h[[8]],  nrow = 3, ncol = 2 )
```
\
\
\
The above grid allows you to see overall trends, the larger individual plots are below.

```{r, echo = FALSE, fig.align = 'center'}
for( i in 1:( length( h ) - 1 ) )
{
    print( h[[i]] )
}
```


\
\
\

## Code

```{r getlabs, results = 'hide', include = FALSE}
## Get all the labels
appLabs <- knitr::all_labels()
appLabs <- setdiff( appLabs, c('getlabs', appLabs[grep("^hide", appLabs)]) )
```

```{r show-code, ref.label = appLabs, echo = TRUE, eval = FALSE}
```
